---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document: default
---

```{r}
library(tidyverse)
library(data.table)
library(combinat)
library(DiagrammeR)
library(webshot)
```

# Designing Process Layouts

```{r include = FALSE, echo = FALSE}
sideX <- 40
sideY <- 40
layoutX <- 2
layoutY <- 2
cells <- (layoutX * layoutY)
dt <- data.frame(
  as.matrix(Matrix::forceSymmetric(matrix(c(0, 10, 40, 30, 
                                            NA, 0, 30, 50, 
                                            NA, NA, 0, 10, 
                                            NA, NA, NA, 0), 4, 4), uplo = "L")))
colnames(dt) <- LETTERS[1:cells]

perms <- lapply(permn(LETTERS[1:cells]), function(x) { matrix(x, layoutY, layoutX) })
let <- c(1:cells)
names(let) <- LETTERS[1:cells]

dist <- unique(permn(c(1:layoutX, 1:layoutY)))
dist <- dist[1:(length(dist)-1)]
totals <- unlist(lapply(perms, function(x) {
  20*dt[let[x[1,1]],let[x[2,1]]] + 
    20*dt[let[x[1,1]],let[x[1,2]]] +
    20*dt[let[x[2,2]],let[x[1,2]]] +
    20*dt[let[x[2,2]],let[x[2,1]]] +
    40*dt[let[x[1,1]],let[x[2,2]]] +
    40*dt[let[x[1,2]],let[x[2,1]]]
}))

min(totals)
perms[which(totals == min(totals))]
```

```{r}
sideX <- 3
sideY <- 2
layoutX <- 3
layoutY <- 2
cells <- (layoutX * layoutY)
dt <- data.frame(
  as.matrix(Matrix::forceSymmetric(matrix(c(0, 30, 0, 34, 50, 25,
                                            NA, 0, 0, 55, 10, 10,
                                            NA, NA, 0, 0, 15, 5,
                                            NA, NA, NA, 0, 0, 0,
                                            NA, NA, NA, NA, 0, 30,
                                            NA, NA, NA, NA, NA, 0), 
                                          cells, cells), uplo = "L")))

colnames(dt) <- LETTERS[1:cells]

perms <- lapply(permn(LETTERS[1:cells]), function(x) {
  matrix(x, layoutY, layoutX)
  })
let <- c(1:cells)
names(let) <- LETTERS[1:cells]

# perms <- list(matrix(c("F", "B", "E", "C", "A", "D"), 2, 3),
#               matrix(c("C", "B", "D", "E", "A", "F"), 2, 3))

pathsLet <- t(combn(names(let), 2))
pathsN <- t(combn(1:cells, 2))

totals <- unlist(lapply(perms, function(x) {
  sum(apply(pathsLet, 1, function(y) {
    start <- which(x == y[1], arr.ind = TRUE)
    end <- which(x == y[2], arr.ind = TRUE)
    ((abs(end[,"row"] - start[,"row"]) * (sideY / layoutY)) +
      (abs(end[,"col"] - start[,"col"]) * (sideX / layoutX))) *
      dt[let[y[1]], y[2]]
  }))
}))

min(totals)
perms[which(totals == min(totals))]
```

# Line Balancing

```{r}
dt <- data.frame("task" = c("A", "B", "C", "D", "E", "E", "F", "G", "H"),
                 "time" = c(55, 30, 35, 22, 50, NA, 15, 5, 10),
                 "predecessor" = c(NA, "A", "A", "B", "B", "C", "C", "F", "G"))
forecastedDemand <- 50
```


## Draw the precedence relationship network

```{r}
writeLines(paste("Total time:", 
                 totalTime <- sum(unique(dt[, c("task", "time")])[, "time"], na.rm = TRUE)))

mermaid(c("graph LR;",
          apply(dt[rowSums(is.na(dt)) == 0,], 1, function(x) { 
            paste0(x["predecessor"], " --> ", x["task"], ";") 
            })
          ))
```

## Potential Output Rate per Hour

```{r}
writeLines(paste("The maximum output per hour is:", 
                 maxOutput <- 3600 / totalTime))

if(forecastedDemand > maxOutput) {
  writeLines(paste("Forecasted demand is greater than maximum output so an assembly line with work stations is a possibility"))
} else {
  writeLines(paste("Forecasted demand is less than maximum output so an assembly line with work stations is not a possibility"))
}
```

## Determine the takt time

```{r}
writeLines(paste("The takt time is:", taktTime <- 3600 / forecastedDemand))
writeLines(paste("The maximum bottleneck output is:", 
                 bottleneckOutput <- 3600 / max(dt$time, na.rm = TRUE)))
```

## Theoretical Minimum Workstations

```{r}
writeLines(paste("The theoretical minimum workstations is:", 
                 minimumStations <- totalTime / taktTime))

if(minimumStations %% 1 != 0) {
  writeLines(paste("However, because you cannot have a partial workstation, the real number is:", ceiling(minimumStations)))
}
```

## Assign tasks to work stations

```{r, echo=FALSE, eval=FALSE, include=FALSE}
stations <- data.frame("station" = 1:4,
                       "eligible" = NA_character_,
                       "selected" = NA_character_,
                       "time" = NA_integer_,
                       "idle" = NA_integer_)

dt$dependencies <- sapply(1:nrow(dt), function(x) { 
  p <- unique(dt$predecessor[1:x])
  paste(p[!is.na(p)], collapse = "") 
  })
dt

dt <- data.table::data.table(dt)[, groupTime := sum(time, na.rm = TRUE), 
                                 by = dependencies]

dt[, groupID := .GRP, by = predecessor]
dt$completed <- sapply(dt$groupID, function(x) { paste0(LETTERS[1:(x-1)], collapse = "") })

dtCopy <- dt[!duplicated(dt[, .(dependencies, groupTime)])]

dtCopy$groupID <- 0
groupID <- 1

for(x in 1:nrow(dtCopy)) {
  for (y in 1:nrow(dtCopy)) {
    if (sum(dtCopy$groupTime[x:y]) <= taktTime) {
      dtCopy$groupID[x:y] <- groupID
    } else {
      groupID <- groupID + 1
    }
  }
}

dtCopy[, groupID := .GRP, by = groupID]

for(x in 1:nrow(stations)) {
  stations$time[x] <- sum(dtCopy$groupTime[x])
  stations$idle[x] <- taktTime - stations$time[x]
}
```

| Work Station | Eligible Tasks | Task Selected | Time | Idle |
| ------------ | -------------- | ------------- | ---- | ---- |
| 1            | A              | A             | 55   | 17   |
| 2            | BC             | BC            | 65   |  7   |
| 3            | DEF            | DE            | 72   |  0   |
| 4            | FGH            | FGH           | 30   | 52   |

```{r}
stations <- data.frame("station" = 1:4,
                       "eligible" = c("A", "BC", "DEF", "FGH"),
                       "selected" = c("A", "BC", "DE", "FGH"),
                       "time" = c(55, 65, 72, 30),
                       "idle" = c(17, 7, 0, 52))
```


## Compute Efficiency and balance delay

```{r}
writeLines(paste("The efficiency is:", sum(stations$time) / (taktTime * ceiling(minimumStations))))
```

# Product Line Balancing Problem

```{r}
dt <- data.frame("task" = c("A", "B", "C", "D", "E", "F", "G", "G"),
                 "time" = c(60, 12, 35, 55, 10, 50, 5, 5),
                 "predecessor" = c(NA, "A", "B", "A", "D", "E", "F", "C"))
forecastedDemand <- 40
```

## 1 Precedence Relationship Network

```{r}
writeLines(paste("Total time:", 
                 totalTime <- sum(unique(dt[, c("task", "time")])[, "time"])))

mermaid(c("graph LR;",
          apply(dt[rowSums(is.na(dt)) == 0,], 1, function(x) { 
            paste0(x["predecessor"], " --> ", x["task"], ";") 
            })
          ))
```

## 2 Determine the Output Rate per hour (day)

```{r}
writeLines(paste("The maximum output per hour is:", 
                 maxOutput <- 3600 / totalTime))
writeLines(paste("The maximum output per 8 hour workday is:", maxOutput * 8))

if(forecastedDemand > maxOutput) {
  writeLines(paste("Forecasted demand is greater than maximum output so an",
                   "assembly line with work stations is a possibility"))
} else {
  writeLines(paste("Forecasted demand is less than maximum output so an", 
                   "assembly line with work stations is not a possibility"))
}
```

## 3 Determine the takt time and maximum bottleneck output

```{r}
writeLines(paste("The takt time is:", taktTime <- 3600 / forecastedDemand))
writeLines(paste("The maximum bottleneck output is:", 
                 bottleneckOutput <- 3600 / max(dt$time)))
```

## 4 Compute the theoretical minimum workstations

```{r}
writeLines(paste("The theoretical minimum workstations is:", 
                 minimumStations <- totalTime / taktTime))

if(minimumStations %% 1 != 0) {
  writeLines(paste("However, because you cannot have a partial workstation,",
                   "the real number is:", 
                   minimumStations <- ceiling(minimumStations)))
}
```

## 5 Assign tasks to workstations

| Work Station | Eligible Tasks | Task Selected | Time | Idle |
| ------------ | -------------- | ------------- | ---- | ---- |
| 1            | AB             | AB            | 72   | 18   |
| 2            | CD             | CD            | 90   | 0    |
| 3            | EFG            | EFG           | 65   | 25   |

```{r}
stations <- data.frame("station" = 1:ceiling(minimumStations),
                       "eligible" = c("AB", "CD", "EFG"),
                       "selected" = c("AB", "CD", "EFG"),
                       "time" = c(72, 90, 65),
                       "idle" = c(18, 0, 25))
```

## 6 Compute Efficiency and balance delay

```{r}
writeLines(paste0("The efficiency is: ", 
                  efficiency <- totalTime / 
                    (taktTime * minimumStations) * 100, "%"))
writeLines(paste0("The balance delay is: ", 100 - efficiency, "%"))
```

