---
title: "MAN 4558 ~ Exam 3"
author: "Gus Lipkin ~ glipkin6737@floridapoly.edu"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(tidyverse)
library(data.table)
```

# Question 1

```{r}
dt <- data.table("Sample" = 1:10,
                 "Obs1" = c(5:8, 3, 9, 5, 4, 5, 2),
                 "Obs2" = c(7, 6, 7, rep(5, 4), 3, 8, 2),
                 "Obs3" = c(6, 9, 5, 6, 7, 2, 4, 2, 8, 5))
dt

valueTable <- data.table("SampleSize" = 2:7,
                         "A2" = c(1.88, 1.02, .73, .58, .48, .42),
                         "D3" = c(rep(0, 5), .08),
                         "D4" = c(3.27, 2.57, 2.28, 2.11, 2.00, 1.92),
                         "d2" = c(1.13, 1.69, 2.06, 2.33, 2.53, 2.70))

valueTable <- valueTable[ncol(dt) - 2]
```

## A. 
> Find the x and R of each sample and record above.

```{r}
dt[, c("xBar", "R") := .(rowMeans(dt[,2:ncol(dt)]),
                          do.call(pmax, dt[,2:ncol(dt)]) - 
                           do.call(pmin, dt[,2:ncol(dt)]))]
dt
```

## B
> Find and record the grand mean $\bar{\bar{x}}$ and $\bar{R}$ in the appropriate boxes.

```{r}
writeLines(paste("The grand mean is:", xBar <- mean(dt$xBar)))
writeLines(paste("The R is:", R <- mean(dt$R)))
```

## C
> Find the UCL and LCL for the $\bar{x}$ chart.

```{r}
writeLines(paste("The UCL for the x chart is:", xBarUCL <- xBar + (valueTable$A2 * R)))
writeLines(paste("The LCL for the x chart is:", xBarLCL <- xBar - (valueTable$A2 * R)))
```

## D
> Find the UCL and LCL for the R chart.

```{r}
writeLines(paste("The UCL for the R chart is:", RUCL <- valueTable$D4 * R))
writeLines(paste("The LCL for the R chart is:", RLCL <- valueTable$D3 * R))
```

## E
> Sketch out a control chart for the x-bar chart. Is the process in control in terms of the “average” vacuum pressure? ______ Yes or No
> If it's continuous

```{r}
dt %>%
  ggplot() +
  geom_hline(yintercept = xBarLCL, color = "red") +
  geom_hline(yintercept = xBar, color = "green") +
  geom_hline(yintercept = xBarUCL, color = "red") +
  geom_line(aes(x = Sample, y = xBar)) +
  geom_point(aes(x = Sample, y = xBar), size = 2) +
  scale_x_continuous("Sample", 1:nrow(dt)) + 
  ylim(c(min(xBarLCL - (xBarLCL * .05), dt$xBar - (dt$xBar * .05)), 
         max(xBarUCL * 1.05, dt$xBar * 1.05)))

if(min(dt$xBar) > xBarLCL & max(dt$xBar) < xBarUCL) {
  writeLines("The process is in control in terms of the average.")
} else {
  writeLines("The process is not in control in terms of the average.")
}
```

## F
> Sketch out a control chart for the R chart. Is the process in control in terms of the “variation” in the process? _______ Yes or No

```{r}
dt %>%
  ggplot() +
  geom_hline(yintercept = R, color = "green") +
  geom_hline(yintercept = RLCL, color = "red") +
  geom_hline(yintercept = RUCL, color = "red") +
  geom_line(aes(x = Sample, y = R)) +
  geom_point(aes(x = Sample, y = R), size = 2) +
  scale_x_continuous("Sample", 1:nrow(dt)) + 
  ylim(c(min(RLCL - (RLCL * .05), dt$R - (dt$R * .05)), 
         max(RUCL * 1.05, dt$R * 1.05)))

if(min(dt$R) > RLCL & max(dt$R) < RUCL) {
  writeLines("The process is in control in terms of the variation")
} else {
  writeLines("The process is not in control in terms of the variation")
}
```

## G
> Find the process standard deviation ($\sigma$)

```{r}
writeLines(paste("The standard deviation is:", sigma <- (R / valueTable$d2)))
```

## H
> Find the appropriate process capability index for the process if the lower and upper specification limits are 8 to 20.

```{r}
LSL <- 2
USL <- 8
if((LSL + USL) / 2 == xBar) { 
  writeLines("The process is centered")
  C <- (USL - LSL) / (6 * sigma) 
} else { 
  writeLines("The process is not centered")
  C <- min((USL - xBar) / (3 * sigma), 
           (xBar - LSL) / (3 * sigma))
}
writeLines(paste("C is:", C))
```


## I
> What is your interpretation of the capability index?

```{r}
if(C < 1) {
  writeLines("Because C is less than one, the process is not capable and process improvement should begin immediately.")
} else {
  writeLines("Because C is greater than one, the process is capabale.")
  if(C < 1.33) {
    writeLines("However, because C is still under 1.33, we should probably improve the process a bit more.")
  }
}
```


# Question 2
```{r}
df <- data.table("towns" = LETTERS[1:4],
                 "distanceA" = c(21, 16, 10.5, 4.5),
                 "distanceB" = c(12, 16, 30, 14),
                 "loads" = c(15, 11, 12, 18))
writeLines(paste("Distance to Martin:", sum(df$loads * df$distanceA)))
writeLines(paste("Distance to Dyersburg:", sum(df$loads * df$distanceB)))
```

The supply center should be located in Martin because the load distance is less than that if it was located in Dyersburg.
