---
title: "Boosting"
output: html_notebook
---

```{r}
library(gbm)
library(ISLR2)
library(caret)
set.seed(1234)
data<-Boston
head(data)
```

Partition the data
```{r}
intrain1<-createDataPartition(y=data$medv, p=0.8, list=FALSE)
train_data1 <- data[intrain1, ]
test_data1 <- data[-intrain1, ]
```

Set up Crossvalidation

```{r}
reg_control <-trainControl(method="cv", number = 10)
```

##Boosting

We will use the gbm package and the Boston dataset to apply boosting

gbm has four tuning parameters

- **n.trees** (# of Boosting Iterations)
- **interaction.depth** (Max Tree Depth i.e. the number of splits in each tree d)
- **shrinkage parameter**  (typically 0.01 or 0.001 and if nothing specified gbm uses a default of 0.001)
- **n.minobsinnode**  (Minimum Terminal Node Size)

```{r}
set.seed(1234)
boost_model <- gbm(medv ~ ., train_data1, distribution = "gaussian", n.trees = 5000, interaction.depth = 4)
summary(boost_model)
```
Based on the relative influence obtained above, we can plot the marginal effects of the most influential variables using _partial dependence plots_ (PDPs). You can learn more about PDPs [here](http://dpmartin42.github.io/posts/r/partial-dependence) and [here](https://christophm.github.io/interpretable-ml-book/pdp.html)

```{r}
plot(boost_model, i = "rm")
plot(boost_model, i = "lstat")
```

```{r}

```
Let us use the test data for prediction

```{r}
yhat_boost <- predict(boost_model, test_data1, n.trees = 5000)
plot(yhat_boost, test_data1$medv)
abline(0, 1)
```

```{r}
mean((yhat_boost - test_data1$medv)^2)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

