---
title: "Regression Trees"
output: html_notebook
---

## Using Tree Package 

```{r}
library(tree) # To construct classification and regression trees
library(ISLR2)
library(tidyverse)
library(data.table)
library(caret)
```

```{r}
dt <- Boston
head(dt)
```

Split the data into training and test data 

```{r}
set.seed(1234)
rowPicker <- sample(c(TRUE, FALSE), nrow(dt), replace = TRUE, prob = c(.5, .5))
train <- dt[rowPicker, ]
test <- dt[!rowPicker, ]
```

Create a regression tree using the training data
```{r}
tr <- tree(medv ~ ., train)
summary(tr)
```
 The above output indicates that only three variables have been used in constructing the tree. Let us plot the tree
 
```{r}
plot(tr)
text(tr, pretty = 0)
```
The variable lstat measures the percentage of individuals with lower socioeconomic status. rm corresponds to the average number of rooms. 
The tree indicates that larger values of rm or lower values of lstat correspond to more expensive houses. 

```{r}
tr_prune <- cv.tree(tr, K = 10)
plot(tr_prune$size, tr_prune$dev, type = "b")
```

```{r}
tr_dPrune <- prune.tree(tr, best = 4)
plot(tr_dPrune)
text(tr_dPrune, pretty = 0)
```

We use the unpruned tree to make predictions on the test dataset

```{r}
reg_yhat <- predict(tr, test)
plot(reg_yhat, test[, 13])
abline(0, 1)
```

Compute the test MSE

```{r}
mean((reg_yhat - test[, 13])^2)
```

## Using the caret package and rpart method

```{r}
library(rpart.plot)
```

Partition the data
```{r}
set.seed(1234)
rowPicker <- sample(c(TRUE, FALSE), nrow(dt), replace = TRUE, prob = c(.8, .2))
train <- dt[rowPicker, ]
test <- dt[!rowPicker, ]
```

Set up Crossvalidation

```{r}
reg_control <- trainControl(method = "cv", number = 10)
```

Model a regression tree

```{r}
tr <- train(medv ~ ., train, method = "rpart", trControl = reg_control)
tr
```

```{r}

```
Plot the pruned tree

```{r}
rpart.plot(tr$finalModel)
```

```{r}

```
Find the MSE on the test data

```{r}
reg_yhat <- predict(tr, test)
plot(reg_yhat, test[, 13])
abline(0, 1)
```


```{r}
mean((reg_yhat - test[, 13])^2)
```

