---
title: "SVM"
output: html_notebook
---

We will use a problem from the back of the book to demonstrate SVM. We will use the Auto dataset to predict whether a given car gets high or low gas mileage. 
(a) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.
(b) Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results. Note you will need to fit the classifier without the gas mileage variable to produce sensible results.
(c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.
(d) Make some plots to back up your assertions in (b) and (c).

```{r}
# Load the required libraries
library(ISLR2)
library(caret)
library(MASS)
library(MLeval)
library(tidyverse)
library(data.table)
set.seed(1234)
```

```{r}
dt <- Auto %>% select(-name)
```
Create a new dataframe to remove missing values and delete the column name

(a) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.

```{r}
dt$mpg <- as.factor(ifelse(dt$mpg > median(dt$mpg), "YES", "NO"))
```

(b) Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results. Note you will need to fit the classifier without the gas mileage variable to produce sensible results.

We will use the caret package to divide the data into testing and training sets
```{r}
rowPicker <- createDataPartition(y=dt$mpg, p=0.8, list=FALSE)
train <- dt[rowPicker,]
test <- dt[-rowPicker,]
```

Initialize the trainControl

```{r}
control <- trainControl(method = "cv", number = 10, classProbs = TRUE, 
                        summaryFunction = twoClassSummary, 
                        savePredictions = TRUE)
```

Create a grid of vales for the tuning parameter C

```{r}
tGrid <- expand.grid(C = c(.01, .05, .1, .25, .5, .75, 1, 1.25, 5, 10))
```

Estimate the SVM Model

```{r}
svmModel <- train(mpg ~ ., train, method = "svmLinear", tuneGrid = tGrid,
                  trControl = control, metric = "ROC", 
                  preProcess = c("center", "scale"))
svmModel
```
Check performance of the linear model

```{r}
svmPred <- predict(svmModel, test)
confusionMatrix(svmPred, test$mpg)
```



(c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.

create a range of values for sigma and C 
```{r}
rGrid <- expand.grid(C = c(.1, .25, .5, .75, 1, 1.25, 1.5),
                     sigma = c(1e-3, 1e-2, 1e-1))
```

Model with Radial Kernel

```{r}
radModel <- train(mpg ~ ., train, method = "svmRadial", tuneGrid = rGrid,
                  trControl = control, metric = "ROC", 
                  preProcess = c("center", "scale"))
radModel
```
Check performance of the radial kernel model

```{r}
radPred <- predict(radModel, test)
confusionMatrix(radPred, test$mpg)
```

Model with polynomial Kernel

```{r}
polyModel <- train(mpg ~ ., train, method = "svmPoly",
                  trControl = control, metric = "ROC", 
                  preProcess = c("center", "scale"))
polyModel
```

Check performance of the polynomial model

```{r}
polyPred <- predict(polyModel, test)
confusionMatrix(polyPred, test$mpg)
```
Compute ROC curves to compare the models using MLeval package

```{r}
plot_comp <- evalm(list1 = list(svmModel, radModel, polyModel),
                   gnames = c("Linear SVC", "Radial SVM", "Polynomial SVM"))
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

