---
title: "Homework 6"
author: "Gus Lipkin ~ glipkin6737@floridapoly.edu"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

> Vacuum pumps are being audited to determine if the process is in control and also capable of meeting the customer specifications. Samples of size n=5 are taken each hour of the day and tested with the data on vacuum recorded. The results of 10 consecutive hours of testing are provided.

```{r}
library(tidyverse)
library(data.table)
```


```{r}
dt <- fread("Homework 6.csv")
dt

valueTable <- data.table("SampleSize" = 2:7,
                         "A2" = c(1.88, 1.02, .73, .58, .48, .42),
                         "D3" = c(rep(0, 5), .08),
                         "D4" = c(3.27, 2.57, 2.28, 2.11, 2.00, 1.92),
                         "d2" = c(1.13, 1.69, 2.06, 2.33, 2.53, 2.70))

valueTable <- valueTable[ncol(dt) - 2]
```

## A. 
> Find the x and R of each sample and record above.

```{r}
dt[, c("xBar", "R") := .(rowMeans(dt[,2:ncol(dt)]),
                          do.call(pmax, dt[,2:ncol(dt)]) - 
                           do.call(pmin, dt[,2:ncol(dt)]))]
dt
```

## B
> Find and record the grand mean $\bar{\bar{x}}$ and $\bar{R}$ in the appropriate boxes.

```{r}
writeLines(paste("The grand mean is:", xBar <- mean(dt$xBar)))
writeLines(paste("The R is:", R <- mean(dt$R)))
```

## C
> Find the UCL and LCL for the $\bar{x}$ chart.

```{r}
writeLines(paste("The UCL for the x chart is:", xBarUCL <- xBar + (valueTable$A2 * R)))
writeLines(paste("The LCL for the x chart is:", xBarLCL <- xBar - (valueTable$A2 * R)))
```

## D
> Find the UCL and LCL for the R chart.

```{r}
writeLines(paste("The UCL for the R chart is:", RUCL <- valueTable$D4 * R))
writeLines(paste("The LCL for the R chart is:", RLCL <- valueTable$D3 * R))
```

## E
> Sketch out a control chart for the x-bar chart. Is the process in control in terms of the “average” vacuum pressure? ______ Yes or No

```{r}
dt %>%
  ggplot() +
  geom_hline(yintercept = xBarLCL, color = "red") +
  geom_hline(yintercept = xBar, color = "green") +
  geom_hline(yintercept = xBarUCL, color = "red") +
  geom_line(aes(x = Sample, y = xBar)) +
  geom_point(aes(x = Sample, y = xBar), size = 2) +
  scale_x_continuous("Sample", 1:nrow(dt)) + 
  ylim(c(xBarLCL - (xBarLCL * .05), xBarUCL * 1.05))

if(min(dt$xBar) > xBarLCL & max(dt$xBar) < xBarUCL) {
  writeLines("The process is in control in terms of the average.")
} else {
  writeLines("The process is not in control in terms of the average.")
}
```

## F
> Sketch out a control chart for the R chart. Is the process in control in terms of the “variation” in the process? _______ Yes or No

```{r}
dt %>%
  ggplot() +
  geom_hline(yintercept = R, color = "green") +
  geom_hline(yintercept = RLCL, color = "red") +
  geom_hline(yintercept = RUCL, color = "red") +
  geom_line(aes(x = Sample, y = R)) +
  geom_point(aes(x = Sample, y = R), size = 2) +
  scale_x_continuous("Sample", 1:nrow(dt)) + 
  ylim(c(RLCL - (RLCL * .05), RUCL * 1.05))

if(min(dt$R) > RLCL & max(dt$R) < RUCL) {
  writeLines("The process is in control in terms of the variation")
} else {
  writeLines("The process is not in control in terms of the variation")
}
```

The process is in control in terms of the average.

## G
> Find the process standard deviation ($\sigma$)

```{r}
writeLines(paste("The standard deviation is:", sigma <- (R / valueTable$d2)))
```

## H
> Find the appropriate process capability index for the process if the lower and upper specification limits are 8 to 20.

```{r}
LSL <- 8
USL <- 20
if((LSL + USL) / 2 == xBar) { 
  writeLines("The process is centered")
  C <- (USL - LSL) / (6 * sigma) 
} else { 
  writeLines("The process is not centered")
  C <- min((USL - xBar) / (3 * sigma), 
           (xBar - LSL) / (3 * sigma))
}
writeLines(paste("C is:", C))
```


## I
> What is your interpretation of the capability index?

```{r}
if(C < 1) {
  writeLines("Because C is less than one, the process is not capable and process improvement should begin immediately.")
} else {
  writeLines("Because C is greater than one, the process is capabale.")
  if(C < 1.33) {
    writeLines("However, because C is still under 1.33, we should probably improve the process a bit more.")
  }
}
```

