---
title: "Assignment 2"
author: "Gus Lipkin ~ glipkin6737@floridapoly.edu"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
  pdf_document: default
---

# 8
> This question involves the use of simple linear regression on the Auto data set.

```{r}
library(data.table)
library(tidyverse)
auto <- fread("../Assignment 1/Auto.csv")
auto$horsepower <- as.numeric(auto$horsepower)
```


## 8a
> Use the `lm()` function to perform a simple linear regression with `mpg` as the response and `horsepower` as the predictor. Use the `summary()` function to print the results. Comment on the output. For example:

```{r}
lm <- lm(mpg ~ horsepower, data = auto)
lm
summary(lm)
```


### 8a i
> Is there a relationship between the predictor and the re- sponse?

Yes. After horsepower is converted to numeric, there is a relationship between `horsepower` and `mpg`.

### 8a ii
> How strong is the relationship between the predictor and the response?

Very strong. The p-value is tiny.

### 8a iii
> Is the relationship between the predictor and the response positive or negative?

Negative. As `horsepower` increases, `mpg` decreases.

### 8a iv
> What is the predicted `mpg` associated with a `horsepower` of 98? What are the associated 95% confidence and prediction intervals?

```{r}
predict(lm, auto)[98]
# confidence interval
predict(lm, auto, interval = "confidence")[98,]
# prediction interval
predict(lm, auto, interval = "prediction")[98,]
```


## 8b
> Plot the response and the predictor. Use the `abline()` function to display the least squares regression line.

```{r}
plot(auto$horsepower, auto$mpg)
abline(lm)
```


## 8c
> Use the `plot()` function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.

```{r}
plot(lm)
```

From the textbook, Figure 3.9: "*Plots of residuals versus predicted (or fitted) values for the Auto data set. In each plot, the red line is a smooth fit to the residuals, intended to make it easier to identify a trend. Left: A linear regression of `mpg` on `horsepower`. A strong pattern in the residuals indicates non-linearity in the data. Right: A linear regression of `mpg` on `horsepower` and `horsepower`. There is little pattern in the residuals.*"

# 9
> This question involves the use of multiple linear regression on the `Auto` data set.

## 9a
> Produce a scatterplot matrix which includes all of the variables in the data set.

```{r}
auto$name <- as.factor(auto$name)
pairs(auto)
```

## 9b
> Compute the matrix of correlations between the variables using the function `cor()`. You will need to exclude the `name` variable, which is qualitative.

```{r}
cor(auto[, -9])
```

## 9c
> Use the `lm()` function to perform a multiple linear regression with `mpg` as the response and all other variables except `name` as the predictors. Use the `summary()` function to print the results. Comment on the output. For instance:

```{r}
lm <- lm(mpg ~ cylinders + displacement + horsepower + weight + acceleration + year + origin, auto)
lm
summary(lm)
```


### 9c i
> Is there a relationship between the predictors and the re- sponse?

There is a relationship between `mpg` and `displacement`, `weight`, `year`, and `origin`.

### 9c ii
> Which predictors appear to have a statistically significant relationship to the response?

There is a relationship between `mpg` and `displacement`, `weight`, `year`, and `origin`.

### 9c iii
> What does the coefficient for the `year` variable suggest?

This suggests that the `mpg` increases alongside the `year`. (Cars become more efficient over time).

## 9d
> Use the `plot()` function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers? Does the leverage plot identify any observations with unusually high leverage?

```{r}
plot(lm)
```

There are a decent amount of outliers among the dataset. There are a few observations with strangely high leverage.

## 9e
> Use the `*` and `:` symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant?

```{r}
summary(lm(mpg ~ cylinders * displacement + horsepower:weight, auto))
```

I guessed and it turns out that my suspicions were correct. `cylinders:diplacement` and `horsepower:weight` were significant.

## 9f
> Try a few different transformations of the variables, such as $log(X)$, $\sqrt{X}$, $X^2$. Comment on your findings.

```{r}
summary(lm(log(mpg) ~ cylinders^2, auto))
summary(lm(log(mpg) ~ cylinders^2 + log(horsepower)^2:sqrt(weight), auto))
```

I'm a little surprised that the regressions didn't break as easily as I thought they would with some basic transformations.

# 10
> This question should be answered using the `Carseats` data set.

```{r}
carseats <- ISLR2::Carseats
```


## 10a
> Fit a multiple regression model to predict `Sales` using `Price`, `Urban`, and `US.`

```{r}
lm <- lm(Sales ~ Price + Urban + US, carseats)
lm
summary(lm)
```

## 10b
> Provide an interpretation of each coefficient in the model. Be careful—some of the variables in the model are qualitative!

As `Price` increases by 1, `Sales` decrease by $.054459$
When `Urban` is `Yes`, `Sales` decrease by $.021916$ (but this isn't significant)
When `US` is `Yes`, `Sales` increases by $1.021916$

## 10c
> Write out the model in equation form, being careful to handle the qualitative variables properly.

$\textrm{Sales}=13.04-.05\textrm{Price}-(.02\textrm{Urban}_{\textrm{Yes}})+(1.20\textrm{US}_{\textrm{Yes}})$

## 10d
> For which of the predictors can you reject the null hypothesis $H0:β_j=0$ ?

`Price` and `USYes` are significant enough with a p-value less than 0.05 so we are 95% confident that they are significant.

## 10e
> On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.

```{r}
lm <- lm(Sales ~ Price + US, carseats)
lm
summary(lm)
```

## 10f
> How well do the models in (a) and (e) fit the data?

The model in (e) fits much better than the one in (a) because it only has significant variables. 

## 10g
> Using the model from (e), obtain $95\%$ confidence intervals for the coefficient(s).

```{r}
confint(lm)
```

## 10h
> Is there evidence of outliers or high leverage observations in the model from (e)?

```{r}
par(mfrow = c(2,2))
plot(lm)
```

There are a few points that look to be outside the normal range. Points 51, 69, and 377 seem to be outliers and points 26, 50, and 368 seem to have high leverage. I'm not worried about them because there are only three in a good-sized dataset. 

# 4
> From the data available at [http://www.stat.ufl.edu/~winner/datasets.html](http://www.stat.ufl.edu/~winner/datasets.html), obtain the data on “Fibre Diameters and Breaking Strenghs for Nextel 610 Fibres.” (please note that there is a typo on the website. It should be Strength and not Strengh) According to the description available there, the expectation is that the log of breaking strength of the fibre should be negatively and linearly related to diameter. (Note log here means natural log if not specified.)

```{r}
dt <- fread("fibre.csv")
colnames(dt) <- c("stren", "diam")
head(dt)
```


## 4a
> Produce a scatter plot of breaking strength against diameter.

```{r}
plot(dt$diam, dt$stren)
```

## 4b
> Produce a scatter plot of the log of breaking strength against diameter. 

```{r}
dt$logStren <- log(dt$stren)
plot(dt$logStren, dt$diam)
```

## 4c 
> Produce a scatter plot of the log of breaking strength against the log of diameter.

```{r}
dt$logDiam <- log(dt$diam)
plot(dt$logStren, dt$logDiam)
```

## 4d 
> Regress breaking strength on diameter. 

```{r}
lm <- lm(stren ~ diam, data = dt)
lm
summary(lm)
```

## 4e
> Regress the log of breaking strength on diameter. 

```{r}
lm <- lm(logStren ~ diam, data = dt)
lm
summary(lm)
```


## 4f 
> Regress the log of breaking strength on the log of diameter. 

```{r}
lm <- lm(logStren ~ logDiam, data = dt)
lm
summary(lm)
```

# 5
> From the data available at [http://www.stat.ufl.edu/~winner/datasets.html](http://www.stat.ufl.edu/~winner/datasets.html), obtain the data on “Variables associated with Permeability and Porosity of Rocks”

```{r}
dt <- fread("rocks.csv")
colnames(dt) <- c("type", "density", "porosity", "logPermeability", "residue", "carbonate", "aGrain", "aSD", "bGrain", "bSD", "calcite", "dolomite")
head(dt)
```


## 5a
> Fit a multiple regression model to predict porosity. Please provide a clean model with only significant variables. Interpret your model results and diagnostics.

```{r}
lm <- lm(porosity ~ ., data = dt)
lmSumm <- summary(lm)
coefs <- coef(lmSumm)[-1,4] <= .05

# This prunes the regression until only significant variables remain
while(length(coefs) != sum(coefs)) {
  form <- paste("porosity ~", paste(names(coefs[coefs]), "+", collapse = " "), collapse = " ")
  form <- substr(form, 1, nchar(form) - 2)
  form
  lm <- lm(formula = form, data = dt)
  lmSumm <- summary(lm)
  coefs <- coef(lmSumm)[-1,4] <= .05
}
lm
lmSumm
par(mfrow = c(2,2))
plot(lm)
```

This is a little silly, but the only significant variable is `logPermeability` and the intercept is even more significant than that is. The $R^2$ value is far from great. I think that because the dataset is rather small, there are more issues that are more prevalent. This is seen in the diagnostic charts where the values and lines are all over the place. Points 5, 8, 10, and 20 seem to be the biggest problems, but that's 13% of the total dataset.

## 5b
> Fit a multiple regression model to predict log(permeability). Please provide a clean model with only significant variables. Interpret your model results and diagnostics.  

```{r}
lm <- lm(logPermeability ~ ., data = dt)
lmSumm <- summary(lm)
coefs <- coef(lmSumm)[-1,4] <= .05

# This prunes the regression until only significant variables remain
while(length(coefs) != sum(coefs)) {
  form <- paste("logPermeability ~", paste(names(coefs[coefs]), "+", collapse = " "), collapse = " ")
  form <- substr(form, 1, nchar(form) - 2)
  form
  lm <- lm(formula = form, data = dt)
  lmSumm <- summary(lm)
  coefs <- coef(lmSumm)[-1,4] <= .05
}
lm
lmSumm
par(mfrow = c(2,2))
plot(lm)
```

This regression is much better, but still not particularly great. The $R^2$ is much better and the Intercept is no longer significant. `Porosity` is significant which is expected based on the results of 5a. There are again a few outliers and leverage points which skew the results a bit.
